---
output:
  github_document:
    keep_md: TRUE
    venue: "gh"
    advertise: FALSE
    session_info: TRUE
    style: TRUE
    comment: "#;-)"
    tidyverse_quiet: FALSE
    std_out_err: TRUE
knit: reprex::reprex_render
---

# Applied Statistics in R
## Natali Tckvitishvili

### Load libraries

```{r, eval = requireNamespace("tidyverse", quietly = TRUE)}
library(tidyverse)
library(ggplot2)
```

### Exercise 1

a. load data & add new variable - good

```{r}
wine <- read.csv("winequality-white.csv", sep = ";")
print(wine)

wine <- mutate(wine, good = ifelse(quality > 5, 1, 0))
```

b. residual.sugar analysis
* histograms for good and bad quality wines

```{r}
ggplot(wine, aes(x = residual.sugar, fill = as.factor(good))) +
geom_histogram(position = "identity", alpha = .5, bins = 100) +
scale_fill_manual(values=c("#69b3a2", "#404080")) +
facet_wrap(vars(good))
```

## Standard output and error

Some output that you see in an interactive session is not actually captured by rmarkdown, when that same code is executed in the context of an `.Rmd` document. When `std_out_err` is `TRUE`, `reprex::reprex_render()` uses a feature of `callr:r()` to capture such output and then injects it into the rendered result.

Look for this output in a special section of the rendered document (and notice that it does not appear right here).

```{r}
system2("echo", args = "Output that would normally be lost")
```

## Session info

Because `session_info` is `TRUE`, the rendered result includes session info, even though no such code is included here in the source document. 
